Model Checking Differentially Private Properties

### 2018-06-07 specifications.tex
polish and spell check

\paragraph{Classical Differential Privacy.}
Consider the survey mechanism (Section~\ref{subsec:survey}). For  $v$
with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_Y}\le
3\myPr{v}{K}{N}{\X \mathit{out}_Y}$ for the probability of satisfying
$\X \mathit{out}_Y$ from $u$ and $v$.
The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $u$
and similarly for $\dpriv{\ln 3}{0} (\X \mathit{out}_N)$. 
Recall that differential privacy requires similar output distributions
on neighbors. The formula $\dpriv{\ln 3}{0} (\X
\mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus
specifies differential privacy for states $+$ and $-$. 
The survey mechanism is $(\ln 3, 0)$-differentially private.

@@

\paragraph{Compositionality.}
Compositionality is one of the building block for differential privacy. For any $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private by the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}. The degradation is rooted in the repeated releases of information. To illustrate this property, we consider the extended survey mechanism which allows two consecutive queries.
In this mechanism, an input is either $+$ or $-$; but 
outputs are $\mathit{out}_Y\mathit{out}_Y$,
$\mathit{out}_Y\mathit{out}_N$, $\mathit{out}_N\mathit{out}_Y$, or
$\mathit{out}_N\mathit{out}_N$. 
The model is depicted in Figure~\ref{figure:2-dp-mc}.
%where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.

\begin{figure}
    \centering
\resizebox{.70\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (p) at  (  -3,  0) { $+$ };
    \node[node] (pY1) at (-1.5, .8) { $s_1$ };
    \node at (-1.5, 1.3) { $\mathit{out}_Y$ };
    \node[node] (pN1) at (-1.5, -.8) { $t_1$ };
    \node at (-1.5, -.3) { $\mathit{out}_N$ };

    \node[node] (Y2) at ( 0,  .8) { $s_2$ };
    \node at (0, 1.3) { $\mathit{out}_Y$ };
    \node[node] (N2) at ( 0, -.8) { $t_2$ };
    \node at (0, -.3) { $\mathit{out}_N$ };

    \node[node] (n) at  (   3,  0) { $-$ };
    \node[node] (nY1) at ( 1.5, .8) { $\bar{s}_1$ };
    \node at (1.5, 1.3) { $\mathit{out}_Y$ };
    \node[node] (nN1) at ( 1.5, -.8) { $\bar{t}_1$ };
    \node at (1.5, -.3) { $\mathit{out}_N$ };

@@

Consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_Y\wedge \X \mathit{out}_Y))$. A path satisfies $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ if the second state satisfies $\mathit{out}_Y$ and the third state satisfies $\mathit{out}_Y$ as well. We verify that this formula is satisfied for states $+$ and $-$. Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively. Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_Y, \mathit{out}_N\}$.

Let us consider two slightly different formulae for comparison:
\begin{itemize}
\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$. In this case we claim
  there is no privacy loss, even though there are two queries. The reason is that the output of the first query is not observed at all. It is easy to verify that it is indeed satisfied by $+$ and $-$.
\item $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$. This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ specifies the one step differential privacy. Observe the inner formula is satisfied by all neighbors. Thus, the outer formula also has no privacy loss.
\end{itemize}

\paragraph{Tighter Privacy Bounds for Composition.}
An advantage of applying model checking is that we may get tighter
bounds for composition. Consider the survey mechanism, and the
property $\dpriv{0}{.5} (\X \mathit{out}_Y)$. Obviously, it holds in
states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_Y} =
\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ in
Figure~\ref{figure:2-dp-mc}. Applying the
compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
$\delta_2=2\delta_1=1$. However, we can check easily that one gets
better privacy parameter $(0,.5)$ using the model checking algorithm
because $\myPr{u}{K}{N}{\mathit{out}_Y} = \frac{9}{16}$,
$\frac{1}{16}$ for $u = +, -$ respectively.
In general, compositional theorems for differential privacy only give
asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
must be calculated carefully and often overly pessimistically. Our
algorithm allows data analysts to choose such parameters more
tightly.

>>>>

\paragraph{Differential Privacy.}
Consider the survey mechanism (Section~\ref{subsec:survey}). For  $v$
with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_1}\le
3\myPr{v}{K}{N}{\X \mathit{out}_1}$ for the probability of satisfying
$\X \mathit{out}_1$ from $u$ and $v$.
The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_1)$ holds in state $u$
and similarly for $\dpriv{\ln 3}{0} (\X \mathit{out}_0)$. 
Recall that differential privacy requires similar output distributions
on neighbors. The formula $\dpriv{\ln 3}{0} (\X
\mathit{out}_1)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_0)$ thus
specifies differential privacy for states $+$ and $-$. 
The survey mechanism is $(\ln 3, 0)$-differentially private.

@@

\paragraph{Compositionality.}
Compositionality is one of the building blocks for differential privacy. For any $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private by the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}. The degradation is rooted in the repeated releases of information. To illustrate this property, we consider the extended survey mechanism which allows two consecutive queries.
In this mechanism, an input is either $+$ or $-$; but 
outputs are $\mathit{out}_1\mathit{out}_1$,
$\mathit{out}_1\mathit{out}_0$, $\mathit{out}_0\mathit{out}_1$, or
$\mathit{out}_0\mathit{out}_0$. 
The model is depicted in Figure~\ref{figure:2-dp-mc}.
%where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.

\begin{figure}
    \centering
\resizebox{.70\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (p) at  (  -3,  0) { $+$ };
    \node[node] (pY1) at (-1.5, .8) { $s_1$ };
    \node at (-1.5, 1.3) { $\mathit{out}_1$ };
    \node[node] (pN1) at (-1.5, -.8) { $t_1$ };
    \node at (-1.5, -.3) { $\mathit{out}_0$ };

    \node[node] (Y2) at ( 0,  .8) { $s_2$ };
    \node at (0, 1.3) { $\mathit{out}_1$ };
    \node[node] (N2) at ( 0, -.8) { $t_2$ };
    \node at (0, -.3) { $\mathit{out}_0$ };

    \node[node] (n) at  (   3,  0) { $-$ };
    \node[node] (nY1) at ( 1.5, .8) { $\bar{s}_1$ };
    \node at (1.5, 1.3) { $\mathit{out}_1$ };
    \node[node] (nN1) at ( 1.5, -.8) { $\bar{t}_1$ };
    \node at (1.5, -.3) { $\mathit{out}_0$ };

@@

Consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_1\wedge \X \mathit{out}_1))$. A path satisfies $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ if the second state satisfies $\mathit{out}_1$ and the third state satisfies $\mathit{out}_1$ as well. We verify that this formula is satisfied for states $+$ and $-$. Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively. Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_1, \mathit{out}_0\}$.

Let us consider two slightly different formulae for comparison:
\begin{itemize}
\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_1)$. In this case we claim
  there is no privacy loss, even though there are two queries. The reason is that the output of the first query is not observed at all. It is easy to verify that it is indeed satisfied by $+$ and $-$.
\item $\dpriv{\ln 3}{0} (\X (\mathit{out}_1\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_1)))$. This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_1)$ specifies the one-step differential privacy. Observe the inner formula is satisfied by all neighbors. Thus, the outer formula also has no privacy loss.
\end{itemize}

\paragraph{Tighter Privacy Bounds for Composition.}
An advantage of applying model checking is that we may get tighter
bounds for composition. Consider the survey mechanism, and the
property $\dpriv{0}{.5} (\X \mathit{out}_1)$. Obviously, it holds in
states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_1} =
\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
(\mathit{out}_1\wedge \X \mathit{out}_1))$ in
Figure~\ref{figure:2-dp-mc}. Applying the
compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
$\delta_2=2\delta_1=1$. However, we can check easily that one gets
better privacy parameter $(0,.5)$ using the model checking algorithm
because $\myPr{u}{K}{N}{\mathit{out}_1} = \frac{9}{16}$,
$\frac{1}{16}$ for $u = +, -$ respectively.
In general, compositional theorems for differential privacy only give
asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
must be calculated carefully and often pessimistically. Our
algorithm allows data analysts to choose better parameters.

### 2018-06-07 subsampling.tex
polish and spell check

The sensitivity of queries is required to apply the (truncated)
$\alpha$-geometric mechanism. Recall that the sensitivity of a query
is the maximal difference of query results on any two neighbors. There are
two problems for mechanisms depending on query sensitivity. First, 
sensitivity of queries can be hard to compute or estimate. Second,
the sensitivity over arbitrary neighbors can be too conservative for
the actual dataset in use. One therefore would like to have
mechanisms independent of query sensitivity.

Subsampling is a technique to design such mechanisms~\cite{DR:14:AFDP}. Concretely, let
us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
is the majority in the dataset while respecting individual
privacy. This can be achieved as
follows~(Algorithm~\ref{algorithm:subsampling-majority}). The
mechanism first samples $m$ 
sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
$d$~(line~\ref{algorithm:subsampling-majority-subsample}). 
It then computes the majority of each sub-dataset and obtains $m$
sub-results. Let $\mathit{count}_R$ and $\mathit{count}_B$ be the
number of sub-datasets with the majority $R$ and $B$
respectively~(line~\ref{algorithm:subsampling-majority-count}). 
Since there are $m$ sub-datasets, we have
$\mathit{count}_R + \mathit{count}_B = m$. To ensure differential
privacy, the mechanism makes sure the difference $| \mathit{count}_R -
\mathit{count}_B |$ is significantly large after perturbation. In
line~\ref{algorithm:subsampling-majority-noise},
$\mathit{Lap}(p)$ denotes a continuous random
variable with the Laplace distribution whose probability density
function is $f(x) = \frac{1}{2p}e^{-|x|/p}$.
If the perturbed difference is sufficiently large,
the mechanism reports the majority of the $m$
sub-results~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,
no information is
revealed~(line~\ref{algorithm:subsampling-majority-mute}).

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SubsamplingMajority}{$d$, $f$}
    \Require{$d \in \{ R, B \}^n$, $f : \{ R, B \}^* \rightarrow \{ R,
      B \}$}
    \State{$q, m \leftarrow 
            \frac{\epsilon}{64\ln (1/\delta)},
            \frac{\log (n/\delta)}{q^2}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \label{algorithm:subsampling-majority-subsample}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \label{algorithm:subsampling-majority-count}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \label{algorithm:subsampling-majority-difference}
    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
    \label{algorithm:subsampling-majority-noise}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
      \label{algorithm:subsampling-majority-output}
    \Else
      \State{\Return {$\bot$}}
      \label{algorithm:subsampling-majority-mute}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling Majority}
  \label{algorithm:subsampling-majority}
\end{algorithm}

Fix the dataset size $n$ and privacy parameters $\epsilon$, $\delta$,
the subsampling majority mechanism can be modeled by a Markov chain.
Figure~\ref{figure:subsampling-majority-markov-chain} gives a sketch
of the Markov chain for $n = 3$. The leftmost four states represent
all possible datasets. Given a dataset, $m$ samples are taken with
replacement. Outcomes of these samples are denoted by
$(\mathit{count}_R, \mathit{count}_B)$. There are only $m + 1$
outcomes from $(m, 0), (m - 1, 1), \ldots, (0, m)$. Each outcome is
represetned by a state in
Figure~\ref{figure:subsampling-majority-markov-chain}. From each
dataset, the probability distribution of all outcomes gives the
transition probability in the figure. Next, observe that  $|
\mathit{count}_R - \mathit{count}_B |$ can have only finitely many
values. The values of $r$ 
(line~\ref{algorithm:subsampling-majority-difference}) hence belong to
a finite set $\{ r_m, \ldots, r_M \}$ with the minimum $r_m$ and
maximum $r_M$. For each $r \in \{ r_m, \ldots,
r_M \}$, it is represented by a state in
Figure~\ref{figure:subsampling-majority-markov-chain}. For instance,
both outcomes $(m, 0)$ and $(0, m)$ transit to the state $r_M = 1/(4q)
- 1$ with probability $1$. For each $r \in \{ r_m, \ldots, r_M \}$,
the probability of having $r + \mathit{Lap}(\frac{1}{\epsilon}) > \ln
(1/\delta)/\epsilon$ (line~\ref{algorithm:subsampling-majority-noise})
is equal to the probability of $\mathit{Lap}(\frac{1}{\epsilon}) > \ln
(1/\delta)/\epsilon - r$. This is equal to $\int^{\infty}_{\ln
  (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|} dx$.
From each state $r \in \{ r_m, \ldots, r_M \}$, it thus goes to the state
$\top$ with probability $\int^{\infty}_{\ln (1/\delta)/\epsilon - r}
\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ and to the state $\bot$ with
probability $1 - \int^{\infty}_{\ln (1/\delta)/\epsilon - r}
\frac{\epsilon}{2}e^{-\epsilon|x|} dx$. Finally, the Markov chain
moves to the state $R$ if $\mathit{count}_R \geq \mathit{count}_B$;
otherwise, it moves to $B$.

@@

      \node[node] (outR) at (5,  1.25) { $R$ };
      \node[node] (outB) at (5, .25) { $B$ };

>>>>

The sensitivity of queries is required to apply the (truncated)
$\alpha$-geometric mechanism. Recall that the sensitivity 
is the maximal difference of query results on any two neighbors. 
Two practical problems may arise for
mechanisms depending on query sensitivity. First, 
sensitivity of queries can be hard to compute. Second,
the sensitivity over arbitrary neighbors can be too conservative for
the actual dataset in use. One therefore would like to have
mechanisms independent of query sensitivity.

Subsampling is a technique to design such mechanisms~\cite{DR:14:AFDP}. Concretely, let
us consider $\mathcal{X} = \{ R, B \}$ (for red and blue team members) and a
dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
is the majority in the dataset while respecting individual
privacy. This can be achieved as
follows~(Algorithm~\ref{algorithm:subsampling-majority}). The
mechanism first samples $m$ 
sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
$d$~(line~\ref{algorithm:subsampling-majority-subsample}). 
It then computes the majority of each sub-dataset and obtains $m$
sub-results. Let $\mathit{count}_R$ and $\mathit{count}_B$ be the
number of sub-datasets with the majority $R$ and $B$
respectively~(line~\ref{algorithm:subsampling-majority-count}). 
Since there are $m$ sub-datasets, we have
$\mathit{count}_R + \mathit{count}_B = m$. To ensure differential
privacy, the mechanism makes sure the difference $| \mathit{count}_R -
\mathit{count}_B |$ is significantly large after perturbation. In
line~\ref{algorithm:subsampling-majority-noise},
$\mathit{Lap}(p)$ denotes a continuous random
variable with the probability density
function $f(x) = \frac{1}{2p}e^{-|x|/p}$ of Laplace distribution.
If the perturbed difference is sufficiently large,
the mechanism reports $1$ if the majority of the $m$
sub-results is
$R$ or $0$ if it is
$B$~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,  no
information is 
revealed~(line~\ref{algorithm:subsampling-majority-mute}).

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SubsamplingMajority}{$d$, $f$}
    \Require{$d \in \{ R, B \}^n$, $f : \{ R, B \}^* \rightarrow \{ R,
      B \}$}
    \State{$q, m \leftarrow 
            \frac{\epsilon}{64\ln (1/\delta)},
            \frac{\log (n/\delta)}{q^2}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \label{algorithm:subsampling-majority-subsample}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \label{algorithm:subsampling-majority-count}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \label{algorithm:subsampling-majority-difference}
    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
    \label{algorithm:subsampling-majority-noise}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $1$ } \textbf{else} \Return {$0$} } 
      \label{algorithm:subsampling-majority-output}
    \Else
      \State{\Return {$\bot$}}
      \label{algorithm:subsampling-majority-mute}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling Majority}
  \label{algorithm:subsampling-majority}
\end{algorithm}

Fix the dataset size $n$ and privacy parameters $\epsilon$, $\delta$,
the subsampling majority mechanism can be modeled by a Markov chain.
Figure~\ref{figure:subsampling-majority-markov-chain} gives a sketch
of the Markov chain for $n = 3$. The leftmost four states represent
all possible datasets. Given a dataset, $m$ samples are taken with
replacement. Outcomes of these samples are denoted by
$(\mathit{count}_R, \mathit{count}_B)$. There are only $m + 1$
outcomes: $(m, 0), (m - 1, 1), \ldots, (0, m)$. Each outcome is
represented by a state in
Figure~\ref{figure:subsampling-majority-markov-chain}. From each
dataset, the probability distribution of all outcomes gives the
transition probability in the figure. Next, observe that  $|
\mathit{count}_R - \mathit{count}_B |$ can have only finitely many
values. The values of $r$ 
(line~\ref{algorithm:subsampling-majority-difference}) hence belong to
a finite set $\{ r_m, \ldots, r_M \}$ with the minimum $r_m$ and
maximum $r_M$. 
%For each $r \in \{ r_m, \ldots, r_M \}$, it is represented by a state
%in Figure~\ref{figure:subsampling-majority-markov-chain}. 
For instance,
both outcomes $(m, 0)$ and $(0, m)$ transit to the state $r_M = 1/(4q)
- 1$ with probability $1$. For each $r \in \{ r_m, \ldots, r_M \}$,
the probability of having $r + \mathit{Lap}(\frac{1}{\epsilon}) > \ln
(1/\delta)/\epsilon$ (line~\ref{algorithm:subsampling-majority-noise})
is equal to the probability of $\mathit{Lap}(\frac{1}{\epsilon}) > \ln
(1/\delta)/\epsilon - r$. This is equal to $\int^{\infty}_{\ln
  (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|} dx$.
From each state $r \in \{ r_m, \ldots, r_M \}$, it hence goes to the state
$\top$ with probability $\int^{\infty}_{\ln (1/\delta)/\epsilon - r}
\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ and to the state $\bot$ with
probability $1 - \int^{\infty}_{\ln (1/\delta)/\epsilon - r}
\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ in
Figure~\ref{figure:subsampling-majority-markov-chain}. Finally, the
Markov chain 
moves to the state $1$ if $\mathit{count}_R \geq \mathit{count}_B$;
otherwise, it moves to $0$. Two dataset states are neighbors if they
differ at most one member. For example, $rrb$ is a neighbor
of $rrr$ and $rbb$ but not $bbb$.

@@

      \node[node] (outR) at (5,  1.25) { $1$ };
      \node[node] (outB) at (5, .25) { $0$ };

### 2018-06-07 syntax.tex
polish and spell check

The syntax of $\dpCTLstar$ state and path formulae is given by:
\begin{eqnarray*}
  \Phi & ::= & p \ |\ \neg \Phi \ |\ \Phi \wedge \Phi \ |\
               \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
  \phi & ::= & \Phi \ |\ \neg\phi  \ |\  \phi\wedge\phi  \ |\  \X \phi \ |\ \phi \until \phi
\end{eqnarray*}
A \emph{state} formula $\Phi$ is either  an atomic proposition
$p$, the negation of a state formula, the conjunction of two state
formulae, the \emph{probabilistic} operator $\PJ{J}$ with $J$
an interval in $[0, 1]$ followed by a path formula, or the
\emph{differentially private} operator $\dpriv{\epsilon}{\delta}$
 with two non-negative real numbers $\epsilon$ and $\delta$ followed
 by a path formula. A
\emph{path} formula $\phi$ is simply a linear temporal logic formula, with temporal operator \emph{next}  ($\X$)
and  \emph{until} operator
($\until$) operator enclosed by two path formulae.
We moreover define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$.

>>>>

The syntax of $\dpCTLstar$ state and path formulae is given by:
\begin{eqnarray*}
  \Phi & ::= & p \ |\ \neg \Phi \ |\ \Phi \wedge \Phi \ |\
               \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
  \phi & ::= & \Phi \ |\ \neg\phi  \ |\  \phi\wedge\phi  \ |\  \X \phi \ |\ \phi \until \phi
\end{eqnarray*}
A \emph{state} formula $\Phi$ is either  an atomic proposition
$p$, the negation of a state formula, the conjunction of two state
formulae, the \emph{probabilistic} operator $\PJ{J}$ with $J$
an interval in $[0, 1]$ followed by a path formula, or the
\emph{differentially private} operator $\dpriv{\epsilon}{\delta}$
 with two non-negative real numbers $\epsilon$ and $\delta$ followed
 by a path formula. A
\emph{path} formula $\phi$ is simply a linear temporal logic formula, with temporal operator \emph{next}  ($\X$)
and  \emph{until} operator
($\until$) enclosed by two path formulae.
We moreover define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$.
